---
title: "Correlation Example"
author: "Daloha Rodr&iacute;guez-Molina"
date: "January 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 1, digits = 2)
```

## Correlation example

This is a simple correlation example I made up about a fictional case in IG. The hypothesis states that the larger the number of posts, the larger the number of followers.

#### 1. Generate random data
First, I'll generate 30 random observations for a variable called `n.posts` that stores the number of posts, and 30 matching observations for the number of followers stored in a variable called `n.followers`:

```{r}
set.seed(350)
n.posts <- abs(round(rnorm(n = 30, mean = 350, sd = 350), 0))
set.seed(700)
n.followers <- abs(round(rnorm(n = 30, mean = 350, sd = 350), 0))
cbind(n.posts, n.followers) # show data
```

#### 2. Including Plots
We generate scatterplots for each of our made-up variables:

```{r}
plot(n.posts)
plot(n.followers)
```

#### 3. Perform correlation:
``` {r}
x <- cor.test(n.posts, n.followers)
```
We can observe that the correlation is somewhat weak, because the correlation is roughly $`r round(x$estimate,2)*100`$%, the p-value is $`r round(x$p.value,4)`$, and the 95%CI crosses the null value ($`r round(x$conf.int,2)[1]`$ ; $`r round(x$conf.int,2)[2]`$).

#### 4. Sort in ascending order and repeat:
Now, we want to sort both variables in ascending order to generate an artificial positive correlation. The data we now obtain looks like this:
```{r}
n.posts <- sort(n.posts)
n.followers <- sort(n.followers)
cbind(n.posts, n.followers) # show data
```

The plots look like this:
```{r}
plot(n.posts) # scatterplots
plot(n.followers)
```

And running the correlation test
```{r}
x <- cor.test(n.posts, n.followers) # (highly significant)
```
we now observe that there is a very high and statistically significant positive correlation of $`r round(x$estimate,2)*100`$% (p-value: $`r x$p.value`$, 95%CI: $`r round(x$conf.int,2)[1]`$ ; $`r round(x$conf.int,2)[2]`$).

#### 5. Sort in descending order
We now want to generate an artificial negative correlation by sorting one of the variables (`n.posts`) in descending order:
```{r}
n.posts <- sort(n.posts, decreasing = TRUE)
n.followers <- sort(n.followers)
cbind(n.posts, n.followers) # show data
```

The scatterplots look like this:
```{r}
plot(n.posts) # scatterplots
plot(n.followers)
```

And the correlation test
```{r}
x <- cor.test(n.posts, n.followers) # (highly significant)
```
shows that there is a very strong and statistically significant negative correlation of $`r round(x$estimate,2)*100`$% (p-value: $`r x$p.value`$, 95%CI: $`r round(x$conf.int,2)[1]`$ ; $`r round(x$conf.int,2)[2]`$).

#### 6. Linear regression
Lastly, we regress `n.followers` against `n.posts`:
```{r}
lm.ig <- lm(n.followers ~ n.posts)
summary(lm.ig)
```
