---
title: "Script for Miryam"
author: "Daloha Rodr&iacute;guez-Molina"
date: "October 29, 2015"
output: html_document
---


### (You may skip these two steps):

#### Generating 102 (6*17) random values of mean 5 and sd 1.5
```{r}
set.seed(1234)
x <- rnorm(102, 5, 1.5)
```


#### Building a matrix from x, with 6 rows and 17 columns
``` {r}
data <- matrix(x,nrow=6,ncol=17)
```


###(You may start here):

#### Sorting out the first column
```{r}
sort(data[,1])
```

#### Create a vector to store differences
```{r}
difference <- NULL
```

#### Calculating the differences: abs(x1-x2), and so on, until abs(x5-x6)
```{r}
for(i in 1:(length(data[,1])-1)) {
  difference[i] <- abs(data[,1][i] - data[,1][i+1])
}
```

#### Create a boolean vector to store whether or not the differences are >=1.5 sd of this first column or not
``` {r}
sd_1.5 <- NULL

for(i in 1:length(difference)) {
  if(difference[i] >= sd(data[,1])*1.5) {
    sd_1.5[i] <- TRUE
  } else {
    sd_1.5[i] <- FALSE
  }
}

```

#### Check results:
```{r}
sd(data[,1])*1.5
difference
sd_1.5
```

###### End of Script
